---
title: "Initial environmental impacts of Piney Point wastewater discharge into Tampa Bay, Florida"
output: 
  bookdown::word_document2:
    reference_docx: my_styles.docx
    number_sections: false
always_allow_html: true
bibliography: refs.bib
author: "Marcus W. Beck (mbeck@tbep.org), Gary E. Raulerson (graulerson@tbep.org), Maya C. Burke (mburke@tbep.org), Joe Whalen (jwhalen@tbep.org), Sheila Scolaro (sscolar@tbep.org), Ed T. Sherwood (esherwood@tbep.org)"
urlcolor: blue
csl: marine-pollution-bulletin.csl
link-citations: true
---

```{r setup, echo = F, warning = F, message = F, results = 'hide'}
# figure path, chunk options
knitr::opts_chunk$set(fig.path = 'figs/', warning = F, message = F, echo = F, cache = F, dev.args = list(family = 'serif'), dpi = 300, warning = F,
  fig.process = function(x) {
  x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
  if (file.rename(x, x2)) x2 else x
  })

# # extract bib entries from online
# bib_scrp('manu_draft.Rmd', 'refs.bib')
```

`r paste('Last manuscript build', Sys.time())`

```{r echo = F, cache = F, eval = F}
spelling::spell_check_files('manu_draft.Rmd')
```

# Abstract

From March 30th to April 9th, 2021, 215 million gallons of legacy phosphate mining wastewater from the Piney Point facility were released into Tampa Bay (Florida, USA).  An estimated ~205 tons of total nitrogen were exported to Lower Tampa Bay, exceeding typical annual nitrogen load estimates in a matter of days.  An initial phytoplankton response was observed in samples closest to the discharge site, with chlorophyll concentrations exceeding 50 ug/L.  Macroalgae blooms of cyanobacteria (*Lyngbya* spp.) were observed in May, with biomass estimated at 0.5 kg/m2 at some locations.  Blooms of *Karenia brevis* were observed beginning in May and continuing through July.  Reported fish kills tracked bloom concentrations, prompting local cleanup efforts to remove over 1700 tons of dead fish.  Combined, these observations indicate abnormal conditions in Tampa Bay following release of wastewater from Piney Point, which is supported by comparison to the decades of baseline environmental monitoring data for the region. 

*Key words*: nitrogen, phosphate mining, Tampa Bay, wastewater, water quality

# Introduction

* Nutrient management paradigm for southwest Florida estuaries, historical progress
* Threats/challenges to protecting water quality: nutrient inputs, climate change
* Wastewater byproducts from mining are a global threat to the quality of surface and groundwater sources worldwide [@Hudson11]. 
     * [@Tayibi09] reviews environmental impact of phosphogypsum production, including contaminantion of groundwater/leaching to coastal waters.  
     * [@Burnett01] evaluate radiochemistry of phosphogypsum at Piney Point to evaluate potential for contamination of groundwater sources [pdf](https://www.osti.gov/etdeweb/servlets/purl/20243518)
    * Citations from [@Perez16] to track down: Santos-Cubatao estuarine system (Brazil; Sanders et al., 2013), in the Lebanese coast (El Samad et al., 2014) and in the Gulf of Gabes (Tunisia; El Zrelli et al., 2016)
    * Dillon report about Grand Bay [link](https://www.wrri.msstate.edu/pdf/2016dillon_finalreport.pdf)
* Phosphate mining industry in Florida, then/now
* Piney Point
* Goals/objectives

This paper provides an initial assessment of environmental conditions in Tampa Bay following the release of 215 million gallons of legacy phosphate mining wastewater over a ten-day period.  The goal is to describe the results of monitoring data collected in response to the discharge event to assess deviation of current conditions relative to long-term, seasonal records of water quality, phytoplankton, seagrass, benthic, and nekton datasets available for the region.  We provide a brief overview of the history of the Piney Point facility, including past wastewater releases and impacts observed in Tampa Bay.  A timeline of events in 2021 is also provided, which is supported by the results from response-based monitoring of conditions in and around Piney Point.  The results of this study provide an initial documentation of impacts to the natural resources of Tampa Bay that can be used to inform long-term assessments of acute wastewater discharge events on the environmental quality of the region.  

# Methods

* Historical timeline: https://docs.google.com/document/d/1KqaEAYEG7pdcGpWbUjONfVKMHHB6_kOkKCyncp4JYKA/edit, https://thebradentontimes.com/piney-point-a-retrospective-p6328-158.htm
* Initial response, data collection
* Synthesis and Analysis

# Results

* 2021 timeline
* Historical context/baseline condition
* effluent characteristics
* water-column water quality changes
* red tide impacts
* Seagrass/macroalgae

# Discussion

* Comparison to other locations/past events - Grand Bay, Bishop Harbor, Huelva estuary [@Perez10;@Perez16]
* Analysis limitations: no smoking gun but 2021 is an anomaly, additional info (benthic diversity TBD, nekton diversity TBD, large mammals, etc.)
* Potential long-term impacts TBD
* Current challenges in TB/southwest FL - OTB, seagrass loss, red tide, climate change
* Risk of decline (IRL ex.), regression of past progress

# Acknowledgments

# Figures

# Tables

# References